vegan_analysis

### 2023-07-24 scripts/0_whole_game.qmd
return from australia
secondary sampling scheme, revised manuscript for ncc
sent to meg and tx 2023-07-21b

Needed things

```{r loads}
vlce <- rast("F:/mosaiced/VLCE2.0/LC_Class_HMM_2015_v20_v20.dat")
bcb <- bcmaps::bc_bound() %>%
  vect()

@@

vlce <- crop(vlce, bcb_rast) %>%
  mask(bcb_rast)
```

@@

>>>>

Needed rasters for 1 2 and 3

```{r loads}
vlce_loc <- "F:/mosaiced/VLCE2.0/LC_Class_terr_bc.dat"

if (!file.exists(vlce_loc)) {
  vlce <- rast("F:/mosaiced/VLCE2.0/LC_Class_HMM_2015_v20_v20.dat")
  bcb <- bcmaps::bc_bound() %>%
    vect()

@@

  vlce <- crop(vlce, bcb_rast) %>%
    mask(bcb_rast)

  writeRaster(vlce, vlce_loc, filetype = "envi")
}
vlce <- rast(vlce_loc)
bcb <- bcmaps::bc_bound() %>%
  vect()

bcb_rast <- bcb %>%
  rasterize(y = vlce) %>%
  crop(bcb)
```

@@

dir.create(sample_loc, showWarnings = F)

### 2023-07-24 scripts/4_vegan.qmd
sent to meg and tx 2023-07-21b

```{r data-prep}
bec_sample_loc <- here::here("data", "samples", "bec_extracted.shp")

@@

bec_df <- bec_samples_e %>% 
  st_as_sf() %>%
  st_drop_geometry() %>%
  select(bec:VarDHI) %>%
  drop_na() %>%
  tibble() %>%
  filter(elev_cv <= 1000) %>%
  filter(elev_p95 >= 2000)

@@

extracted <- to_rda %>%
  select(elev_p95,
         elev_cv,
         # elev_stddev,
         percentage_first_returns_above_2m)

@@

rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + total_biomass, # modelled
      data = to_rda)

@@

for_radar_loads <- all_data_loadings %>%
  filter(!endsWith(var, "DHI")) %>%
  mutate(axis_loading = ifelse(sig == "NS", 0, abs(axis_loading))) %>%
  select(var, axis, axis_loading) %>%
  pivot_wider(names_from = var, values_from = axis_loading) %>%
  mutate(cluster = paste(axis, "- All Data")) %>%
  select(-axis) %>%
  relocate(cluster)

var_out <- varpart(dhi_tab, extracted, modelled)

var_out %>% 
  plot(Xnames = c("Derived", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()
```

@@

# vlce sampling regime to get the samples up
vlce_samples <- read_sf(here::here("data", "samples", "vlce_extracted.shp"))

vlce_df <- vlce_samples %>%
  st_drop_geometry() %>%
  select(-category_1, -category_2) %>%
  rename(category = strata) %>%
  relocate(bec) %>%
  drop_na() %>%
  ungroup()

names(vlce_df) <- names_manual

vlce_normal <- vlce_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x)))

to_nest <- vlce_normal %>%
  select(-percentage_first_returns_above_mean,
         -loreys_height,
         -elev_stddev,
         -elev_mean)

class_nested <- to_nest %>%
  group_by(category) %>%
  slice_sample(n = 3000) %>%
  nest() %>% 
  ungroup() %>%
  rename(strata = category)

@@

  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + total_biomass, # modelled, put condition here for a conditional rda
      data = tibble)
}

@@

  extracted <- tibble %>%
    select(elev_p95,
           elev_cv,
           # elev_stddev,
           percentage_first_returns_above_2m)

@@

  varpart(dhi_tab, extracted, modelled)
}
```

@@

fracts_out <- outs %>%
  mutate(indfracts = map(var, .f = function(x) {
    x$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
      pivot_wider(names_from = type, values_from = Adj.R.squared)
    }
    )) %>%
  unnest(indfracts) %>%
  arrange(desc(residual)) %>%
  select(strata, extracted, modelled, overlap) %>%
  pivot_longer(c(extracted, modelled, overlap)) %>%
  mutate(value = ifelse(value > 0, value, 0)) %>%
  pivot_wider()

fracts_out %>% 
  mutate(total = extracted + modelled + overlap) %>%
  arrange(strata %in% forest_types,
          desc(total))
```

```{r}
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "All Data") %>% 
  relocate(strata)

@@

fracts_tot <- fracts_all %>% 
  mutate(total = extracted + modelled + overlap) %>%
  select(strata, total)

fracts_bar_tab <- fracts_all %>%
  pivot_longer(-strata) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "All Data", "All", group)) %>%
  left_join(fracts_tot) %>%
  mutate(zone_nm = ifelse(is.na(zone_nm), strata, zone_nm))

@@

fracts_bar_plot <- fracts_bar_tab %>%
  ggplot(aes(x = value, y = fct_reorder(zone_nm, total), fill = fct_rev(fct_relevel(str_to_title(name), "Overlap", after = 1)))) +
  geom_col(position = "stack") +
  facet_grid(rows = vars(fct_reorder(group, total, .desc = T)),
             scales = "free_y",
             space = "free") +
  labs(x = "Variance Explained",
       y = NULL,
       fill = NULL) +
  theme(legend.position = "bottom",
        panel.spacing.y = unit(0, "lines")) +
  guides(fill = guide_legend(reverse = T)) +
  scale_x_continuous(labels = scales::label_percent())

fracts_bar_plot
ggsave(here::here("outputs", "fracts_bar_plot.png"), plot = fracts_bar_plot)
```

```{r raincloud-plot}
library(ggdist)
fracts_out %>%
  mutate(total = extracted + modelled + overlap) %>%
  #left_join(keys$bec, by = c("bec" = "zone")) %>%
  pivot_longer(-strata) %>%
  mutate(name = str_to_title(name)) %>%
  mutate(value = ifelse(value < 0, 0, value),
         class = ifelse(strata %in% forest_types, 
                        "Forest Types", 
                        "BEC Zones")) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye(
    # adjust bandwidth
    adjust = .5,
    # move to the right
    justification = -0.2,
    # remove the slub interval
    .width = 0,
    point_colour = NA
    ) +
  geom_boxplot(
    width = 0.12,
    # removing outliers
    outlier.color = NA,
    alpha = 0.5
  ) +
  stat_dots(
    # ploting on left side
    side = "left",
    # adjusting position
    justification = 1.1,
    # adjust grouping (binning) of observations
    binwidth = 0.0025,
    #position = "stack",
    # aes(fill = group, col = group),
    alpha = 0.5,
    aes(shape = class)
  ) +
  labs(y = NULL,
       x = "% Variance Explained",
       fill = "BEC Group",
       col = "BEC Group") +
  scale_x_continuous(labels = scales::label_percent()) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, .2))

@@

```{r path-setup}
path_base <- tibble(x1 = c(0,
             0,
             0,
             0,
             1),
       y1 = c(1,
              -1,
              1,
              -1,
              0),
       x2 = c(1.85,
              1.85,
              .85,
              .85,
              1.85),
       y2 = c(.15,
              -.15,
              .15,
              -.15,
              0)) %>%
  ggplot(aes(x = x1, y = y1)) +
  geom_segment(aes(xend = x2, yend = y2), arrow = arrow()) +
  geom_point(data = tibble(x1 = c(.85, .85), 
                           y1 = c(-.5, .5)),
             size = 60,
             shape = 21,
             col = "grey") +
  # geom_text(aes(x = 0, y = -1.15), label = "Modelled\nAttributes") +
  # geom_text(aes(x = 0, y = 1.15), label = "Extracted\nAttributes") +
  # geom_text(aes(x = .65, y = 0), label = "Overlap") +
  # geom_text(aes(x = 2, y = 0), label = "Dynamic\nHabitat\nIndices") +
  theme_void() +
  scale_x_continuous(limits = c(-0.1, 2.1))

@@

path_r2_plot <- tibble(type = c("extracted", "modelled", "overlap"),
       x1 = c(0.85, 0.85, 0.85),
       y1 = c(0.75, -0.75, 0))
```

@@

var_out <- varpart(dhi_tab %>%
                     select(any_of(var_oi)),
                   extracted, modelled)

# plot(var_out, Xnames = c("Extracted", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()

@@

clean_loads <- sig_loads %>%
  filter(anova_sig != "NS",
         loading_sig != "NS") %>%
  bind_rows(all_data_loadings %>%
              mutate(strata = "All Data") %>%
              rename(loading_sig = sig)) %>%
  select(strata:axis_loading) %>%
  mutate(axis_loading = abs(axis_loading))
```

##### run to here for analysis to be done
radar chart based on groups, not hierarchical clustering

@@

radar_loads <- clean_loads %>%
  pivot_wider(names_from = var, 
              values_from = axis_loading) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "All Data", "All Data", group),
         group = ifelse(axis == "RDA2", "RDA2", group)) %>%
  group_by(group) %>%
  summarize(across(CumDHI:elev_p95, \(x) mean(x, na.rm = T))) %>%
  mutate(group = fct_relevel(group, "RDA2", after = Inf),
         group = fct_relevel(group, "Forest Types", after = 1))

radar_labels <- radar_loads %>%
  select(percentage_first_returns_above_2m:elev_p95) %>%
  colnames() %>%
  tibble(variable = .) %>%
  left_join(keys$continuous) %>%
  pull(var_long)

@@

boxplots <- clean_loads %>% 
  left_join(keys$continuous, by = c("var" = "variable")) %>%
  mutate(var_long = ifelse(is.na(var_long), var, var_long)) %>%
  select(!(divide_by:unit)) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(axis == "RDA2", "RDA2", group),
         group = ifelse(strata == "All Data", "All Data", group),
         inout = ifelse(endsWith(var, "DHI"), "Response", "Predictor")) %>%
  mutate(group = fct_relevel(group, "RDA2", after = Inf),
         group = fct_relevel(group, "Forest Types", after = 1)) %>%
  ggplot(aes(x = abs(axis_loading), y = fct_rev(fct_reorder(var_long, abs(axis_loading))))) +
  geom_boxplot(outlier.colour = "#00000000", width = .5) +
  geom_jitter(height = 0.1, aes(col = group, shape = group)) +
  facet_grid(cols = vars(axis),
             rows = vars(inout),
             scales = "free_y",
             space = "free_y") +
  labs(x = "Axis Loading (absolute)",
       y = NULL,
       col = NULL,
       shape = NULL) +
  theme(panel.spacing.y = unit(0, "lines"),
        legend.position = "right") +
  scale_shape_manual(values = c(17, 15, 16, 16, 16, 16, 16, 16)) +
  scale_colour_manual(values = cols)

@@

ggsave(here::here("outputs", "radar_boxplot.png"),
       plot = radar_boxplot,
       device = "png")
```

@@

# list to iterate
stratum <- clean_loads %>%
  pull(strata) %>%
  unique()

# make sure everything gets change in rda fig
possibles <- clean_loads %>%
  select(var, axis) %>%
  complete() %>%
  mutate(possible = paste(var, axis, sep = "-")) %>%
  pull(possible) %>%
  unique()

# fractions dfs
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "All Data") %>% 
  relocate(strata)

@@

all_varexp <- rda_eig %>%
  filter(name != "RDA3") %>%
  mutate(strata = "All Data") %>%
  rename(axis = name) %>%
  select(-value)

@@

for (s in stratum) {
  strata_tp <- clean_loads %>%
    filter(strata == s) %>%
    mutate(textreplace = paste(var, axis, sep = "-"),
           axis_loading = as.character(round(axis_loading, digits = 2)))

@@

  for(p in possibles) {
    strata_drawio <- str_replace(strata_drawio, p, "NS")
  }

  strata_fract <- fracts_all %>%
    filter(strata == s) %>%
    pivot_longer(extracted:overlap) %>%
    mutate(value = (value * 100) %>%
             round(1) %>%
             paste0(., "%"))

@@

  for(i in 1:length(fract_names)) {
    strata_drawio <- str_replace(strata_drawio, fract_names[[i]], fract_value[[i]])
  }

@@

  for(i in 1:length(axis_names)) {
    strata_drawio <- str_replace(strata_drawio, axis_names[[i]], axis_vals[[i]])
  }

@@

```{r}
working <- rgb_loads_split[[1]]

make_rgb_map <- function(tibble) {

  rgb_cols = tibble %>%
    pull(value) %>%
    sort()

  bec_dissolve %>%
    left_join(tibble, by = c("zone" = "strata")) %>%
    mutate(ifelse(is.na(value), "grey", value)) %>%
    ggplot() +
    geom_sf(aes(geometry = geometry, fill = value), colour = NA) +
    scale_fill_manual(values = rgb_cols) +
    theme_void() +
    theme(legend.position = "none")
}

ggsave(here::here("outputs", "rgbplots_norm.png"), patchwork_rgbplots,
       height = 6,
       width = 8)

print("png saved")

ggsave(here::here("outputs", "rgbplots_norm.pdf"), patchwork_rgbplots,
       height = 6,
       width = 8)

print("pdf saved")
```

@@

>>>>

```{r data-prep}
bec_sample_loc <- here::here("data", "samples", "vlce_extracted.shp")

@@

bec_df <- bec_samples_e %>% 
  st_as_sf() %>%
  st_drop_geometry() %>%
  select(bec:VarDHI) %>%
  drop_na() %>%
  tibble() %>%
  filter(elev_cv <= 1000) %>% #erroneous pixels
  filter(elev_p95 >= 2000) # non-treed pixels

@@

primary <- to_rda %>%
  select(elev_p95,
         elev_cv,
         # elev_stddev,
         percentage_first_returns_above_2m)

@@

rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/primary
      basal_area + gross_stem_volume + total_biomass, # modelled
      data = to_rda)

@@

for_radar_loads <- all_data_loadings %>%
  filter(!endsWith(var, "DHI")) %>%
  mutate(axis_loading = ifelse(sig == "NS", 0, abs(axis_loading))) %>%
  select(var, axis, axis_loading) %>%
  pivot_wider(names_from = var, values_from = axis_loading) %>%
  mutate(cluster = paste(axis, "- British Columbia")) %>%
  select(-axis) %>%
  relocate(cluster)

var_out <- varpart(dhi_tab, primary, modelled)

var_out %>% 
  plot(Xnames = c("Primary", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("primary", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()
```

@@

class_nested <- to_nest %>%
  group_by(category) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = category)

# # vlce sampling regime to get the samples up
# switched away from this to a single double stratified sample
# vlce_samples <- read_sf(here::here("data", "samples", "vlce_extracted.shp"))
# 
# vlce_df <- vlce_samples %>%
#   st_drop_geometry() %>%
#   select(-category_1, -category_2) %>%
#   rename(category = strata) %>%
#   relocate(bec) %>%
#   drop_na() %>%
#   ungroup()
# 
# names(vlce_df) <- names_manual
# 
# vlce_normal <- vlce_df %>%
#   mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x))) 
# 
# to_nest <- vlce_normal %>%
#   select(-percentage_first_returns_above_mean,
#          -loreys_height,
#          -elev_stddev,
#          -elev_mean)
# 
# 
# 
# 
# class_nested <- to_nest %>%
#   group_by(category) %>%
#   slice_sample(n = 3000) %>%
#   nest() %>% 
#   ungroup() %>%
#   rename(strata = category)

@@

  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/primary
      basal_area + gross_stem_volume + total_biomass, # modelled, put condition here for a conditional rda
      data = tibble)
}

@@

  primary <- tibble %>%
    select(elev_p95,
           elev_cv,
           # elev_stddev,
           percentage_first_returns_above_2m)

@@

  varpart(dhi_tab, primary, modelled)
}
```

@@

fracts_out <- outs %>%
  mutate(indfracts = map(var, .f = function(x) {
    x$part$indfract %>%
      mutate(type = c("primary", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
      pivot_wider(names_from = type, values_from = Adj.R.squared)
    }
    )) %>%
  unnest(indfracts) %>%
  arrange(desc(residual)) %>%
  select(strata, primary, modelled, overlap) %>%
  pivot_longer(c(primary, modelled, overlap)) %>%
  mutate(value = ifelse(value > 0, value, 0)) %>%
  pivot_wider()

fracts_out %>% 
  mutate(total = primary + modelled + overlap) %>%
  arrange(strata %in% forest_types,
          desc(total))
```

```{r}
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "British Columbia") %>% 
  relocate(strata)

@@

fracts_tot <- fracts_all %>% 
  mutate(total = primary + modelled + overlap) %>%
  select(strata, total)

fracts_bar_tab <- fracts_all %>%
  pivot_longer(-strata) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "British Columbia", "BC", group)) %>%
  left_join(fracts_tot) %>%
  mutate(zone_nm = ifelse(is.na(zone_nm), strata, zone_nm))

@@

fracts_bar_plot <- fracts_bar_tab %>%
  ggplot(aes(x = value, y = fct_reorder(zone_nm, total), fill = str_to_title(name))) +
  geom_col(position = "stack") +
  facet_grid(rows = vars(fct_reorder(group, total, .desc = T)),
             scales = "free_y",
             space = "free") +
  labs(x = "Variance Explained",
       y = NULL,
       fill = NULL) +
  theme(legend.position = "bottom",
        panel.spacing.y = unit(0, "lines")) +
  guides(fill = guide_legend(reverse = T)) +
  scale_x_continuous(labels = scales::label_percent())

fracts_bar_plot
ggsave(here::here("outputs", "fracts_bar_plot.png"), plot = fracts_bar_plot,
       height = 6, width = 8)
```

```{r raincloud-plot}
library(ggdist)
fracts_out %>%
  mutate(total = primary + modelled + overlap) %>%
  #left_join(keys$bec, by = c("bec" = "zone")) %>%
  pivot_longer(-strata) %>%
  mutate(name = str_to_title(name)) %>%
  mutate(value = ifelse(value < 0, 0, value),
         class = ifelse(strata %in% forest_types, 
                        "Forest Types", 
                        "BEC Zones")) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye(
    # adjust bandwidth
    adjust = .5,
    # move to the right
    justification = -0.2,
    # remove the slub interval
    .width = 0,
    point_colour = NA
    ) +
  geom_boxplot(
    width = 0.12,
    # removing outliers
    outlier.color = NA,
    alpha = 0.5
  ) +
  stat_dots(
    # ploting on left side
    side = "left",
    # adjusting position
    justification = 1.1,
    # adjust grouping (binning) of observations
    binwidth = 0.0025,
    #position = "stack",
    # aes(fill = group, col = group),
    alpha = 0.5,
    aes(shape = class)
  ) +
  labs(y = NULL,
       x = "% Variance Explained",
       fill = "BEC Group",
       col = "BEC Group") +
  scale_x_continuous(labels = scales::label_percent()) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, .2))

@@

```{r path-setup}
path_base <- tibble(x1 = c(0,
             0,
             0,
             0,
             1),
       y1 = c(1,
              -1,
              1,
              -1,
              0),
       x2 = c(1.85,
              1.85,
              .85,
              .85,
              1.85),
       y2 = c(.15,
              -.15,
              .15,
              -.15,
              0)) %>%
  ggplot(aes(x = x1, y = y1)) +
  geom_segment(aes(xend = x2, yend = y2), arrow = arrow()) +
  geom_point(data = tibble(x1 = c(.85, .85), 
                           y1 = c(-.5, .5)),
             size = 60,
             shape = 21,
             col = "grey") +
  # geom_text(aes(x = 0, y = -1.15), label = "Modelled\nAttributes") +
  # geom_text(aes(x = 0, y = 1.15), label = "primary\nAttributes") +
  # geom_text(aes(x = .65, y = 0), label = "Overlap") +
  # geom_text(aes(x = 2, y = 0), label = "Dynamic\nHabitat\nIndices") +
  theme_void() +
  scale_x_continuous(limits = c(-0.1, 2.1))

@@

path_r2_plot <- tibble(type = c("primary", "modelled", "overlap"),
       x1 = c(0.85, 0.85, 0.85),
       y1 = c(0.75, -0.75, 0))
```

@@

var_out <- varpart(dhi_tab %>%
                     select(any_of(var_oi)),
                   primary, modelled)

# plot(var_out, Xnames = c("Primary", "Modelled"))

path_r2 <- var_out$part$indfract %>%
      mutate(type = c("primary", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
  tibble()

@@

clean_loads_raw <- sig_loads %>%
  filter(anova_sig != "NS",
         loading_sig != "NS") %>%
  bind_rows(all_data_loadings %>%
              mutate(strata = "British Columbia") %>%
              rename(loading_sig = sig)) %>%
  select(strata:axis_loading)

clean_loads_abs <- clean_loads_raw %>%
  mutate(axis_loading = abs(axis_loading))
```

##### run to here for analysis to be done

radar chart based on groups, not hierarchical clustering

@@

radar_loads <- clean_loads %>%
  pivot_wider(names_from = var, 
              values_from = axis_loading) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "British Columbia", "British Columbia", group),
         group = ifelse(axis == "RDA2", "RDA2", group)) %>%
  group_by(group) %>%
  summarize(across(CumDHI:elev_p95, \(x) mean(x, na.rm = T))) %>%
  mutate(group = fct_relevel(group, "RDA2", after = Inf),
         group = fct_relevel(group, "Forest Types", after = 1))

radar_labels <- radar_loads %>%
  select(-group, -ends_with("DHI")) %>%
  colnames() %>%
  tibble(variable = .) %>%
  left_join(keys$continuous) %>%
  pull(var_long)

@@

boxplots <- clean_loads %>% 
  left_join(keys$continuous, by = c("var" = "variable")) %>%
  mutate(var_long = ifelse(is.na(var_long), var, var_long)) %>%
  select(!(divide_by:unit)) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(axis == "RDA2", "RDA2", group),
         group = ifelse(strata == "British Columbia", "British Columbia", group),
         inout = ifelse(endsWith(var, "DHI"), "Response", "Predictor")) %>%
  mutate(group = fct_relevel(group, "RDA2", after = Inf),
         group = fct_relevel(group, "Forest Types", after = 1)) %>%
  ggplot(aes(x = abs(axis_loading), y = fct_rev(fct_reorder(var_long, abs(axis_loading))))) +
  geom_boxplot(outlier.colour = "#00000000", width = .5) +
  geom_jitter(height = 0.1, aes(col = group, shape = group)) +
  facet_grid(cols = vars(axis),
             rows = vars(inout),
             scales = "free_y",
             space = "free_y") +
  labs(x = "Axis Loading (absolute)",
       y = NULL,
       col = NULL,
       shape = NULL) +
  theme(panel.spacing.y = unit(0, "lines"),
        legend.position = "right") +
  scale_shape_manual(values = c(17, 15, 16, 16, 16, 16, 16, 16)) +
  scale_colour_manual(values = cols)

@@

ggsave(here::here("outputs", "radar_boxplot.png"),
       plot = radar_boxplot,
       device = "png",
       height = 6,
       width = 8)
```

# drawio

@@

# list to iterate
stratum <- clean_loads_raw %>%
  pull(strata) %>%
  unique()

# make sure everything gets change in rda fig
possibles <- clean_loads_raw %>%
  select(var, axis) %>%
  complete() %>%
  mutate(possible = paste(var, axis, sep = "-")) %>%
  pull(possible) %>%
  unique()

# fractions dfs
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "British Columbia") %>% 
  relocate(strata)

@@

all_varexp <- rda_eig %>%
  filter(name != "RDA3") %>%
  mutate(strata = "British Columbia") %>%
  rename(axis = name) %>%
  select(-value)

@@

for (s in stratum) {
  print(s)
  strata_tp <- clean_loads_raw %>%
    filter(strata == s) %>%
    mutate(textreplace = paste(var, axis, sep = "-"),
           axis_loading = as.character(round(axis_loading, digits = 2)))

@@

  for (p in possibles) {
    strata_drawio <- str_replace(strata_drawio, p, "NS")
  }

  strata_fract <- fracts_all %>%
    filter(strata == s) %>%
    pivot_longer(primary:overlap) %>%
    mutate(value = (value * 100) %>%
             round(1) %>%
             paste0(., "%"))

@@

  for (i in 1:length(fract_names)) {
    strata_drawio <-
      str_replace(strata_drawio, fract_names[[i]], fract_value[[i]])
  }

@@

  for (i in 1:length(axis_names)) {
    strata_drawio <-
      str_replace(strata_drawio, axis_names[[i]], axis_vals[[i]])
  }

@@

@@

# new radar loads

```{r}
clean_loads %>%
  pivot_wider(names_from = var, 
              values_from = axis_loading) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), strata, group),
         group = ifelse(strata == "British Columbia", "British Columbia", group)) %>%
  group_by(group, axis) %>%
  summarize(across(CumDHI:elev_p95, \(x) mean(x, na.rm = T))) %>%
  pivot_longer(CumDHI:elev_p95) %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  left_join(keys$continuous, by = c("name" = "variable")) %>%
  mutate(var_long = ifelse(is.na(var_long), name, var_long),
         order = endsWith(var_long, "DHI")) %>%
  ggplot(aes(x = fct_reorder(var_long, value, .desc = T), y = value)) +
  geom_col() +
  #coord_polar() +
  coord_flip() +
  facet_grid(cols = vars(group), rows = vars(axis)) +
  labs(x = "Variable",
       y = "Axis Loading (absolute)")

clean_loads %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), strata, group),
         group = ifelse(strata == "British Columbia", "British Columbia", group)) %>%
  ggplot(aes(x = var, y = axis_loading, col = group)) +
  geom_line()
```

```{r}
clean_loads %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), strata, group),
         group = ifelse(strata == "British Columbia", "British Columbia", group)) %>%
  pivot_wider(names_from = var, values_from = axis_loading) %>%
  select(-zone_nm, -order) %>%
  replace(is.na(.), 0) %>%
  GGally::ggparcoord(columns = 4:12,
                     groupColumn = "group",
                     scale = "globalminmax") +
  facet_wrap(~axis) +
  coord_flip()
```

# parallel coordinate plots

```{r}
data <- clean_loads_abs %>%
  left_join(keys$bec %>%
              select(strata = zone, group)) %>%
  left_join(keys$continuous %>%
              select(var = variable, var_long)) %>%
  mutate(group = ifelse(is.na(group), strata, group),
         var_long = ifelse(is.na(var_long), var, var_long)) %>%
  group_by(group, var_long, axis) %>%
  summarize(axis_loading = mean(axis_loading)) %>%
  ungroup() %>%
  complete(group, var_long, axis, fill = list(axis_loading = NA)) %>%
  mutate(inout = ifelse(endsWith(var_long, "DHI"), "Response", "Predictor"), 
         group = fct_relevel(group, "British Columbia", 
                             "Broadleaf", 
                             "Coniferous", 
                             "Mixed Wood", 
                             "Wetland-Treed", 
                             after = 0),
         var_long = fct_relevel(var_long, 
                                c("Canopy Height", "Canopy Cover", "Structural Complexity",
                                "Basal Area", "Aboveground Biomass", "Gross Stem Volume",
                                "Cumulative DHI", "Variation DHI", "Minimum DHI")))

# nojitter/lines
parcoord_noflip <- data %>%
  ggplot(aes(x = var_long, y = axis_loading)) +
  #geom_boxplot(col = "#00000025", outlier.alpha = 0) +

  geom_line(aes(col = group, group = group), lty = "dashed") +
  geom_point(aes(col = group, group = group, shape = group, fill = group)) +
  # geom_hline(yintercept = 0, lty = "dashed", col = "red") +
  facet_wrap(~paste(axis, inout),
             scale = "free") +
  #theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL,
       y = "Axis Loading (absolute)",
       col = NULL,
       shape = NULL,
       fill = NULL) +
  # coord_flip() +
  scale_colour_discrete_c4a_cat(palette = "bold") +
  scale_fill_discrete_c4a_cat(palette = "bold") +
  scale_shape_manual(values = c(21, 22, 22, 22, 22, 25, 25, 25, 25, 25))

parcoord_noflip

ggsave(here::here("outputs", "parcoord_noflip.png"), plot = parcoord_noflip,
       height = 6, width = 8)

ggsave(here::here("outputs", "parcoord_noflip.svg"), plot = parcoord_noflip,
       height = 6, width = 8)
```

```{r}
make_rgb_map <- function(tibble) {

  input <- bec_dissolve %>%
    left_join(tibble, by = c("zone" = "strata")) %>%
    mutate(value = ifelse(is.na(value), "#000000", value))

  rgb_cols = input %>%
    st_drop_geometry() %>%
    pull(value) %>%
    unique() %>%
    sort()

  input %>%
    ggplot(aes(x = zone, y = 2, fill = value)) +
    geom_col() +
    scale_fill_manual(values = rgb_cols)

   input %>%
    ggplot() +
    geom_sf(aes(geometry = geometry, fill = value, col = value), linewidth = 0.1) +
    scale_fill_manual(values = rgb_cols) +
    scale_color_manual(values = rgb_cols) +
    theme_void() +
    theme(legend.position = "none")
}

labels = c("DHI", "B - Modelled RDA1", "A - Primary RDA1", "DHI", "D - Modelled RDA2", "C - Primary RDA2")

# for (i in 1:length(rgb_maps)) {
#   print(i)
#   gg <- rgb_maps[[i]] + labs(title = labels[[i]])
#   
#   ggsave(here::here("outputs", "black_plots", paste(i, ".png")),
#          plot = gg,
#          height = 3,
#          width = 3)
# }

ggsave(here::here("outputs", "test_black.png"), rgb_maps[[6]])

cowplot_rgb <- cowplot::plot_grid(rgb_maps[[3]], rgb_maps[[2]],
                                  rgb_maps[[6]], rgb_maps[[5]],
                                  labels = c("A - Primary RDA1", "B - Modelled RDA1",
                                             "C - Primary RDA2", "D - Modelled RDA2"),
                                  vjust = 1.25)

ggsave(here::here("outputs", "rgbplots_cow.png"), cowplot_rgb, height = 7, width = 8, bg = "white")

# ggsave(here::here("outputs", "rgbplots_norm.png"), patchwork_rgbplots,
#        height = 6,
#        width = 8)
# 
# print("png saved")

# ggsave(here::here("outputs", "rgbplots_norm.pdf"), patchwork_rgbplots,
#        height = 6,
#        width = 8)
# 
# print("pdf saved")
```
